---
title: "How to we handle the categorical predictors?"
author: Bicko, Jonathan & Ben
date: "2021 Nov 01 (Mon)"
compact-title: false
output:
  pdf_document:
    extra_dependencies: ["float"]
---


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.pos = "H"
	, out.extra = ""
	, message=FALSE
	, warning=FALSE
	, echo=FALSE
	, comment=""
)

library(shellpipes)
library(dplyr)
library(ggplot2)
library(vareffects); varefftheme()
library(emmeans)
library(effects)

commandEnvironments()
makeGraphics()
 
set.seed(2121)

load("variable_predictions_funs.rda")
```

\section{Single categorical, single continuous predictors}

```{r global_vars}
N <- 100
```

```{r single_cat_single_cont}
single_cat_single_cont_sim <- linearsim(nHH=N
	, form=~1+x1+x2
	, betas=NULL
	, pgausian=list(p=1)
	, pcat=list(p=1
		, fun=sample
		, nlevels=3
		, labels=c("Christian","Muslim","Jews")
		, prob=c(0.4, 0.35, 0.25)
		, replace=TRUE
	)
	, link_scale=TRUE
	, vnames=c("age", "rel", "hhsize")
)
single_cat_single_cont_sim_df <- single_cat_single_cont_sim$data
head(single_cat_single_cont_sim_df)
```

```{r single_cat_single_cont_mod}
single_cat_single_cont_mod <- lm(hhsize ~ age + rel, data=single_cat_single_cont_sim_df)
```

\subsection{religion}

\subsubsection{varpred}

```{r single_cat_single_cont_vpred1}
single_cat_single_cont_vpred1 <- varpred(single_cat_single_cont_mod
	, "rel"
	, handle.inter="emmeans"
	, returnall=TRUE
)
single_cat_single_cont_vpred1$raw$model.matrix
```

\subsubsection{emmeans}

```{r single_cat_single_cont_em1}
single_cat_single_cont_em1 <- emmeans(single_cat_single_cont_mod
	, specs=~rel
)
single_cat_single_cont_em1@linfct
```

\subsubsection{effects}

```{r single_cat_single_cont_eff1}
single_cat_single_cont_eff1 <- Effect("rel"
	, single_cat_single_cont_mod
)
eff_mm <- single_cat_single_cont_eff1$model.matrix
attr(eff_mm, "contrasts") <- NULL
attr(eff_mm, "assign") <- NULL
eff_mm
```

\subsection{age}

\subsubsection{varpred}

```{r single_cat_single_cont_vpred_age}
quant <- seq(0,1,length.out=4)
age_at <- quantile(single_cat_single_cont_sim_df$age, quant)
single_cat_single_cont_vpred_age <- varpred(single_cat_single_cont_mod
	, "age"
	, at=list(age=age_at)
	, handle.inter="emmeans"
	, returnall=TRUE
)
single_cat_single_cont_vpred_age$raw$model.matrix
```

\subsubsection{emmeans}

```{r single_cat_single_cont_em_age}
single_cat_single_cont_em_age <- emmeans(single_cat_single_cont_mod
	, specs=~age
	, at=list(age=age_at)
	, weights="proportional"
)
single_cat_single_cont_em_age@linfct
```

\subsubsection{effects}

```{r single_cat_single_cont_eff_age}
single_cat_single_cont_eff_age <- Effect("age"
	, single_cat_single_cont_mod
	, xlevels=list(age=age_at)
)
eff_mm <- single_cat_single_cont_eff_age$model.matrix
attr(eff_mm, "contrasts") <- NULL
attr(eff_mm, "assign") <- NULL
eff_mm
```

\section{age and religion interaction}

```{r single_cat_single_cont_mod_inter}
single_cat_single_cont_mod_inter <- lm(hhsize ~ age*rel, data=single_cat_single_cont_sim_df)
```

\subsection{religion$\star$age}

\subsubsection{varpred}

```{r single_cat_single_cont_vpred1_inter}
age_mean <- mean(single_cat_single_cont_sim_df$age)
single_cat_single_cont_vpred1_inter <- varpred(single_cat_single_cont_mod_inter
	, c("rel", "age")
	, x.var="rel"
	, at=list(age=age_at)
	, handle.inter="emmeans"
	, returnall=TRUE
)
single_cat_single_cont_vpred1_inter$raw$model.matrix
```

\subsubsection{emmeans}

```{r single_cat_single_cont_em1_inter}
single_cat_single_cont_em1_inter <- emmeans(single_cat_single_cont_mod_inter
	, specs=~rel*age
	, at=list(age=age_at)
)
single_cat_single_cont_em1_inter@linfct
```

\subsubsection{effects}

```{r single_cat_single_cont_eff1_inter}
single_cat_single_cont_eff1_inter <- Effect(c("rel", "age")
	, single_cat_single_cont_mod_inter
	, x.var="rel"
	, xlevels=list(age=age_at)
)
eff_mm <- single_cat_single_cont_eff1_inter$model.matrix
attr(eff_mm, "contrasts") <- NULL
attr(eff_mm, "assign") <- NULL
eff_mm
```


\subsection{Ignore interaction}

\subsection{religion}

\subsubsection{varpred}

```{r single_cat_single_cont_vpred1_inter_ign}
single_cat_single_cont_vpred1_inter_ign <- varpred(single_cat_single_cont_mod_inter
	, "rel"
	, handle.inter="emmeans"
	, returnall=TRUE
)
single_cat_single_cont_vpred1_inter_ign$raw$model.matrix
```


\subsubsection{emmeans}

```{r single_cat_single_cont_em1_inter_ign}
single_cat_single_cont_em1_inter_ign <- emmeans(single_cat_single_cont_mod_inter
	, specs=~rel
	, nesting=NULL
)
single_cat_single_cont_em1_inter_ign@linfct
```

\subsubsection{effects}

```{r single_cat_single_cont_eff1_inter_ign}
single_cat_single_cont_eff1_inter_ign <- Effect("rel"
	, single_cat_single_cont_mod_inter
)
eff_mm <- single_cat_single_cont_eff1_inter_ign$model.matrix
attr(eff_mm, "contrasts") <- NULL
attr(eff_mm, "assign") <- NULL
eff_mm
```

\subsection{age}

\subsubsection{varpred}

```{r single_cat_single_cont_vpred_age_inter_ign}
single_cat_single_cont_vpred_age_inter_ign <- varpred(single_cat_single_cont_mod_inter
	, "age"
	, at=list(age=age_at)
	, handle.inter="emmeans"
	, returnall=TRUE
)
single_cat_single_cont_vpred_age_inter_ign$raw$model.matrix
```

\subsubsection{emmeans}

```{r single_cat_single_cont_em_age_inter_ign}
single_cat_single_cont_em_age_inter_ign <- emmeans(single_cat_single_cont_mod_inter
	, specs=~age
#	, cov.keep="age"
	, at=list(age=age_at)
	, nesting=NULL
	, weights="proportional"
)
single_cat_single_cont_em_age_inter_ign@linfct
```

\subsubsection{effects}

```{r single_cat_single_cont_eff_age_inter_ign}
single_cat_single_cont_eff_age_inter_ign <- Effect("age"
	, single_cat_single_cont_mod_inter
	, xlevels=list(age=age_at)
)
eff_mm <- single_cat_single_cont_eff_age_inter_ign$model.matrix
attr(eff_mm, "contrasts") <- NULL
attr(eff_mm, "assign") <- NULL
eff_mm
```


\section{Two categorical, single continuous predictors}

```{r two_cat_single_cont}
two_cat_sim <- linearsim(nHH=N
	, form=~1+x1+x2+x3
	, betas=NULL
	, pgausian=list(p=1)
	, pcat=list(p=2
		, fun=sample
		, nlevels=c(3,2)
		, labels=list(c("Christian","Muslim","Jews"), c("Male", "Female"))
		, prob=list(c(0.4, 0.35, 0.25), c(0.6, 0.4))
		, replace=TRUE
	)
	, link_scale=TRUE
	, vnames=c("age", "rel", "gender", "hhsize")
)
two_cat_sim_df <- two_cat_sim$data
age_at <- quantile(two_cat_sim_df$age, quant)
head(two_cat_sim_df)
```

```{r two_cat_mod}
two_cat_mod <- lm(hhsize~rel*gender+age, data=two_cat_sim_df)
```

\subsection{religion$\star$gender}

\subsubsection{varpred}

```{r two_cat_vpred_inter}
two_cat_vpred_inter <- varpred(two_cat_mod
	, c("rel", "gender")
	, x.var="rel"
	, returnall=TRUE
)
two_cat_vpred_inter$raw$model.matrix
```

\subsubsection{emmeans}

```{r two_cat_em_inter}
two_cat_em_inter <- emmeans(two_cat_mod
	, specs=~rel*gender
	, weights="proportional"
)
two_cat_em_inter@linfct
```

\subsubsection{effects}

```{r two_cat_eff_inter}
two_cat_eff_inter <- Effect(c("rel", "gender")
	, two_cat_mod
	, x.var="rel"
)
eff_mm <- two_cat_eff_inter$model.matrix
attr(eff_mm, "contrasts") <- NULL
attr(eff_mm, "assign") <- NULL
eff_mm
```

\subsection{age}

\subsubsection{varpred}

```{r two_cat_vpred_inter_age}
two_cat_vpred_inter_age <- varpred(two_cat_mod
	, "age"
	, at=list(age=age_at)
	, returnall=TRUE
)
two_cat_vpred_inter_age$raw$model.matrix
```

\subsubsection{emmeans}

```{r two_cat_em_inter_age}
two_cat_em_inter_age <- emmeans(two_cat_mod
	, specs=~age
	, at=list(age=age_at)
	, weights="proportional"
	, nesting="rel %in% gender"
)
two_cat_em_inter_age@linfct
```

\subsubsection{effects}

```{r two_cat_eff_inter_age}
two_cat_eff_inter_age <- Effect("age"
	, two_cat_mod
	, xlevels=list(age=age_at)
)
eff_mm <- two_cat_eff_inter_age$model.matrix
attr(eff_mm, "contrasts") <- NULL
attr(eff_mm, "assign") <- NULL
eff_mm
```

\subsection{religion}

\subsubsection{varpred}

```{r two_cat_vpred_inter_gender}
two_cat_vpred_inter_rel <- varpred(two_cat_mod
	, "rel"
	, returnall=TRUE
)
two_cat_vpred_inter_rel$raw$model.matrix
```

\subsubsection{emmeans}

```{r two_cat_em_inter_rel}
two_cat_em_inter_rel <- emmeans(two_cat_mod
	, specs=~rel
	, weights="proportional"
	, nesting=NULL
)
two_cat_em_inter_rel@linfct
```



**Check**

```{r}
colMeans(model.matrix(two_cat_mod))
```

