---
title: "Focal interaction for multi-parameter input variables"
author: Bicko, Jonathan & Ben
date: "2023 Mar 08 (Wed)"
compact-title: false
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE}
library(shellpipes)
startGraphics(height=4)

library(dplyr)
library(ggplot2); theme_set(theme_bw(base_size=16))

## Binning function
binfun <- function(mod, focal, bins=50, groups=NULL, ...) {
	if (!is.null(groups)) {
		bins_all <- c(groups, "bin")
	} else {
		bins_all <- "bin"
	}
	if (!inherits(mod, "data.frame")) {
		mf <- model.frame(mod)
	} else {
		mf <- mod
	}
	check_df <- (mf
		%>% arrange_at(focal)
		%>% mutate(bin=ceiling(row_number()*bins/nrow(.)))
		%>% group_by_at(bins_all)
		%>% summarise_all(mean)
		%>% mutate(model="binned")
	)
	return(check_df)
}



df <- rdsRead()
```

The aim is to use a simple example to show what `varpred` does (there is more to that though in the package code) in the case of multi-parameter focal interaction.

Consider a model with the focal multi-parameter:

\[
mass = \beta_0 + \beta_1{nitro} + \beta_2{nitro}^2 + \beta_3{phos} + \epsilon
\]

```{r model}
mod <- lm(mass ~ poly(nitro, 2) + phos, df)
summary(mod)
```

Picking _nitro_ as the focal predictor, the non-focal predictor (_phos_) is set at its model center reference value. For the _focal_ predictor, we simply choose representative values from the input variable. Thus the central prediction is:

\[
\hat{mass} =\hat{\beta_0} + \hat{\beta_1}{nitro} + \hat{\beta_2}{nitro}^2 + \hat{\beta_3}{\bar{phos}}
\]

Let us try to do this the base R way:

```{r baseR}
focal <- df$nitro

beta <- coef(mod)
mod_mat <- model.matrix(mod)

non_focal_mean <- mean(mod_mat[, "phos"])

## Set non-focal model center reference value
mod_mat[, "phos"] <- non_focal_mean
head(mod_mat)

## Generate central predictions
base_pred <- as.vector(mod_mat %*% beta)
```

This averages to the observed average biomass:

```{r compare_baseR}
base <- c(observed=mean(df$mass), base=mean(base_pred))
print(base)
```

But the curve (which averages to the observed) does not pass through this point (black dot)

```{r base_plot}
df$base_pred <- base_pred
base_plot <- (ggplot(df, aes(x=nitro, y=base_pred))
	+ geom_line()
	+ geom_point(aes(y=mass), colour="grey")
	+ geom_hline(yintercept=base[["observed"]], lty=2, colour="blue")
	+ geom_hline(yintercept=base[["base"]], lty=3, colour="red")
	+ geom_point(data=data.frame(mass=base[["observed"]], nitro=mean(focal))
		, aes(x=nitro, y=mass)
		, size=4
		, colour="black"
	)
)
print(base_plot)
```

Let us add binned observations (dark red dots)

```{r binned}
binned_df <-  binfun(df, "nitro", bins=8)
base_plot <- (base_plot
	+ geom_point(data=binned_df, aes(x=nitro, y=mass), size=2, colour="darkred")
)
print(base_plot)
```

**Summary**

- The reference point for the non-focal predictor is chosen at the model center
- All the predictor variables corresponding to the multi-parameter focal variable are evaluated using the chosen focal values
- The average of the predictions is identical to the average of the observed biomass but the curve does not pass through the point (assuming the center of the multi-parameter focal variable is the mean of nitro -- in real sense we don't know the this.)
- These are based on simple base R implementation but agrees with what varpred does.
- **What are we missing in dealing with the multi-parameter focal variables?**
